A Reimplementation of Python’s dict using Simple Tabulation Hashing and Linear Probing

### 2012-05-20 docs/paper.tex
added a new page for appendix
Fixes

The current CPython dictionary implementation uses an iterative XORed polynomial
terms for calculating the hash of the strings. In this implementation the hash
value for sequential strings differ only in the least significant bits, for
instance: hash("6.851") = 7824665118634166871 and hash("6.852") =
7824665118634166868. This behavior gives good results for sequential strings
which is a common case. When the table size is $2^i$ taking the least
significant $i$ bits as the index can populate the table without a lot of
collisions. This approach gives a better than random results in this case and is
also simple and fast to compute.

@@

Based on this new method the probe sequence will depend on nearly all the bits
of a hash value. We should note that the value of  perturb shift (currently 5)
is crucial for a good probing behavior. Since it should be small enough so that
all the significant bits have an  effect on the probe sequence but at the same
time it should be large enough so that in really bad cases the high-order hash
bits can affect the earlier iterations.

@@

We also experimented with indexing using every pair of bytes in the string and
using $2^16$ tabulation table entries.

@@

\bibitem{python} Python Software Foundation \emph{},
\texttt{http://python.org/} 2012.

\bibitem{python} Nick Welch. Hash-table-shootout. 
\texttt{https://github.com/mackstann/hash-table-shootout} 2012.

\bibitem{python} Nicholas Piël. Benchmark of Python WSGI Servers.

@@

\end{thebibliography}
\appendix
\section{Tornado Application}
Sample Tornado 2.2.1 application used for benchmark:
 \begin{verbatim}
import tornado.httpserver
import tornado.ioloop
import tornado.wsgi

>>>>

The current CPython dictionary implementation uses an iterative XORed polynomial
terms for calculating the hash of the strings. In this implementation the hash
value for sequential strings differ only in the least significant bits, for
instance: hash("6.851") = 7824665118634166871 and hash("6.852") =
7824665118634166868. This behavior gives good results for sequential strings
which is a common case. When the table size is $2^i$ taking the least
significant $i$ bits as the index can populate the table without a lot of
collisions. This approach gives better than random results in this case and is
also simple and fast to compute.

@@

Based on this new method the probe sequence will depend on nearly all the bits
of a hash value. We should note that the value of  perturb shift (currently 5)
is crucial for a good probing behavior.  It should be small enough so that
all the significant bits have an effect on the probe sequence but at the same
time it should be large enough so that in really bad cases the high-order hash
bits can affect the earlier iterations.

@@

We also experimented with indexing using every pair of bytes in the string,
treating them as 16bit integers and
using $2^{16}$ tabulation table entries.

@@

\bibitem{python} Python Software Foundation \emph{},
\texttt{http://python.org} 2012.

\bibitem{hashshoot} Nick Welch. Hash-table-shootout. 
\texttt{https://github.com/mackstann/hash-table-shootout} 2012.

\bibitem{wsgi} Nicholas Piël. Benchmark of Python WSGI Servers.

@@

\end{thebibliography}
\appendix
\newpage
\noindent{\LARGE\bfseries Appendix}
\section{Tornado Application}
Sample Tornado 2.2.1 application used for benchmark:
 \begin{verbatim}
import tornado.httpserver
import tornado.ioloop
import tornado.wsgi
